Exact analytic solution for slope flows with spatially varying eddy viscosity and diffusivity

### 2015-10-25 paper_BLM.tex
old solution

% ABSTRACT
\begin{abstract}
An exact analytic solution of the steady-state Prandtl model equations is derived, valid for spatially varying eddy diffusivities (O'Briens type) and Prandtl number of unity. 
In this formulation profiles show significant variations in both phase and amplitude of minima-maxima with respect to the classic constant eddy viscosity model and the more recent, approximate, WKB solution. 
The near wall region is characterized by a relatively stronger surface inversion and velocity gradients, the low-level-jet is further displaced toward the wall, and its peak velocity strongly depends on the model parameter, suggesting a tighter coupling between dynamics and thermodynamics.
\end{abstract}

\section{Introduction}
Natural convection of turbulent stratified flows over sloping surfaces is ubiquitous in nature and engineering, and is of interest not only as a fundamental problem in itself, but also because of the important role it plays over a broad range of scales and applications. 
On a local scale, for instance, it governs local weather conditions, affecting atmospheric transport processes and dispersion of scalars such as heat and humidity \citep{Monti2002, Nylen2004b}. 
On a regional scale, it drives the development of deep-convective circulations \citep{Egger1985, Parish1998} and is responsible for intense cyclonic vorticity in the middle and upper troposphere \citep{Parish1992}.
Further, recent experimental field campaigns \citep{Chu1987, Oerlemans1993} have shown that persistent katabatic winds characterize the atmospheric boundary layer over snow-ice surfaces and glaciers, and therefore an accurate characterization of such flows is an essential component of understanding and modeling of the weather and climate.
However, the complex dynamics (e.g. occurrence of intermittency, waves, Kelvin-Helmholtz instabilities and low-level jets) and the lack of a satisfactory similarity theory for such flows \citep{Nadeau2012} pose a heavy burden in terms of computational requirements for numerical modellers; in most cases the required resolution is prohibitively costly \citep{Shapiro2004, Shapiro2004a, Fedorovich2009, Fedorovich2009d, Burkholder2011a}.
Because of this, conceptual models are still of great interest, and represent a valid tool for the characterization of such systems. 
A cornerstone in the understanding of natural convection of stratified fluid along a sloping surface is represented by the classic Prandtl analytic model \citep{Prandtl}, and its more recent extensions/generalizations to include the effects of Coriolis force \citep{Gutman1964}, external winds \citep{Lykosov_1972}, surface heterogeneity \citep{Egger1985, Shapiro2007, Oldroyd2014}.
The Prandtl model approximates the atmosphere in a Boussinesq sense and describes a steady flow over a thermally perturbed unbounded planar sloping surface that lies within a stratified environment, given by the following system of equations:
%
\begin{linenomath*}
\begin{eqnarray}
	N^2 u(z) \sin{\alpha} & = & d( K_H db/dz )/dz,
	\label{eqn1u} \\
	%
	b(z) \sin{\alpha} & = & -d( K_M du/dz )/dz,
	\label{eqn1b}
\end{eqnarray}
\end{linenomath*}
%
where $z$ denotes the normal-to-slope coordinate directions, $u$ is the downslope velocity, $b$ is buoyancy, $N$ is the buoyancy frequency characterizing the system (related to the background stratification), $\alpha$ is the slope angle and $K_M$ and $K_H$ denote the eddy viscosity and diffusivity (an eddy viscosity/diffusivity model has been used to parametrize turbulent fluxes of momentum and buoyancy).
Equations are integrated in $z \in \ [0,H]$ with boundary conditions $u(0)=0$, $u(H) = 0$, $b(0)=b_s$ and $b(H) = 0$ ($b_s > 0$ for upslope flows, whereas $b_s<0$ for downslope flows).

The flow is assumed to be invariant in the downslope direction and the model can be used to determine its vertical structure \citep{NAPPO1987}.
Eqs. \ref{eqn1u} and \ref{eqn1b} state that downslope (upslope) convection of cold (warm) air is balanced by momentum flux divergence, and that along-slope generated buoyancy is balanced by buoyancy flux divergence; thus the model is applicable away from ridges and valleys, so that advection terms are negligible \citep{NAPPO1987}.
The Prandtl constant-K model is known to be overdissipative in the near surface regions, and therefore not able to represent the observed strong surface gradients of temperature and momentum that commonly characterize katabatic and anabatic flows \citep{Oerlemans1998, grisogono2001katabatic}. 
Simple variations in the eddy diffusivity profiles were introduced in an exact analytic solution by \citet{gutman_1983}, and more recently \citet{grisogono2001katabatic} considered general variations in the vertical structure of the eddy diffusivities, and derived  solutions based on the WKB approximation.
The WKB solution is able to account for additional dynamics while still retaining an elegant form.
However, that theory is only applicable when the model parameters $(K_M, K_H)$ vary more slowly than the solution $(u,b)$, and the validity of such a constraint for slope flows has been the subject of great debate \citep{Grisogono2002}.
In the following, we derive a closed-form solution to the Prandtl-model equations, valid for O'Brien's type \citep{O'Brien1970} eddy diffusivities and representing an exact alternative to the WKB solution -- for the chosen form of the model parameter.

% ###################################################################
\section{The Analytic Solution}
%
We consider the one-dimensional Prandtl model equations \ref{eqn1u} and \ref{eqn1b}. Assuming $Pr=1$ (i.e. $K_M=K_H=K$) and assigning length, velocity and buoyancy scales as $L=\max{(K)}^{1/2} N^{-1/2} \sin^{-1/2}{\alpha}$, $U=b_s N^{-1}$ and $B=b_s$ Eqs. \ref{eqn1u} and \ref{eqn1b} can be reduced to:
%
\begin{linenomath*}
\begin{subequations}
\label{ode}
\begin{align}
	u^+& = d( K^+ db^+/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{eqn1u_norm} \\
	%
	b^+ & = -d( K^+ du^+/dz^+ )/dz^+, &&\mathrm{in}  \ [0,H^+]
	\label{eqn1b_norm}
\end{align}
\end{subequations}
\end{linenomath*}
%
respectively, where $K^+$ is a normalized eddy viscosity, $z^+=zL^{-1}$, $u^+ = uU^{-1}$, $b^+=b B^{-1}$, with boundary conditions $u^+(0)=0$, $u^+(H^+) = 0$, $b^+(0)=\pm 1$ and $b^+(H^+) = 0$.
%
Further, we introduce the following variables 
%
\begin{linenomath*}
\begin{equation}
   f_1= b^+ - i u^+ \hspace{3mm} \textrm{and} \hspace{3mm} f_2=b^+ + iu^+,
   \label{combinationVars}
\end{equation}
\end{linenomath*}
%
and upon substitution of Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm} into Eqs. \ref{combinationVars}, the system can be decoupled to obtain two (independent) complex ordinary differential equations (ODEs) for the canonical variables $f_1$ and $f_2$:
%
\begin{linenomath*}
\begin{subequations}
\label{ode_norm}
\begin{align}
	i \cdot f_1 &=  d(  K^+ df_1/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{ode1} \\
	i \cdot f_2 &= - d( K^+ df_2/dz^+ )/dz^+, && \mathrm{in}  \ [0,H^+]
	\label{ode2}
\end{align}
\end{subequations}
\end{linenomath*}
%
with boundary conditions $f_i(0)=\pm 1$ and $f_i(H^+) = 0$, $i=1,2$.

We assume eddy diffusivities in line with the classic O'Brien's type \citep{O'Brien1970}, viz. $K^+ = A^+ (z^++\epsilon^+)(z^+-H^+-\epsilon^+)^2$, where
$\epsilon^+$ is an additional parameter that we introduce in order to shift the zeroes of $K^+$ outside the interval of integration $[0,H^+]$.
$\epsilon^+$ can be tuned to provide the desired value of viscosity / diffusivity at the wall.

In order to solve Eqs. \ref{ode1} and \ref{ode2}, we observe that both have the form
%
\begin{linenomath*}
\begin{equation*}
	f= \mp i \cdot d/dz^+ ( K^+ df/dz^+ ),
\end{equation*}
\end{linenomath*}
%
and therefore can be solved using the same method. First, we rewrite the equation in canonical form:
%
\begin{linenomath*}
\begin{align}
	\label{can_form}
	f''+P f'+ Qf &=0, && \mathrm{in}  \ [0,H^+]
\end{align}
\end{linenomath*}
%
where $P(z^+)=(K^+)'/K^+$ and $Q(z^+)=1/(\pm i K^+)$. Second, it is an easy computation to show that, if $f$ solves Eq. \ref{can_form} then $g(z^+):=f( (H^++2 \epsilon^+) z^+- \epsilon^+)$ solves:
%
\begin{linenomath*}
\begin{align}
	\label{can_form_sec}
	g''+\widetilde{P} g'+ \widetilde{Q} g =0, && \mathrm{in}  \ [0,1]
\end{align}
\end{linenomath*}
%
where $\widetilde{P}(z^+)=\gamma'(z^+)/\gamma(z^+)$ and $\widetilde{Q}(z^+)=1/(\pm i \cdot A^+ \cdot (H^++2\epsilon^+) \gamma(z^+))$, with $\gamma(z^+)=z^+(z^+-1)^2$.
Eqn. \ref{can_form_sec} is a second order ODE with three regular singular points at $z^+ = 0,1$ and $\infty$, as in \citet{Morse1954a}. 
This special case is known as the \emph{equation of Papperitz} and its general solution is:
%
%\begin{widetext}
\begin{linenomath*}
\begin{equation}
	g(z^+) = \alpha (1-z^+)^\mu \ _2F_1 (\mu, 1- \mu', 1+\mu-\mu', 1-z^+)
	+\beta (1-z^+)^{\mu'} \ _2F_1 (\mu', 1- \mu, 1+\mu'-\mu, 1-z^+),
	\label{gen_sol_g}
\end{equation}
\end{linenomath*}
%\end{widetext}
%
where $\ _2F_1$ are \emph{Gaussian hypergeometric functions} and $\mu$, $\mu'$ are the solutions to the degree two equation
%
\begin{linenomath*}
\begin{equation*}
	x^2+x+1/(\pm i A^+(H^++2 \epsilon^+))=0,
\end{equation*}
\end{linenomath*}
%
depending on which of Eqs. \ref{can_form_sec} we are solving.
Note that the general solution to Eq. \ref{can_form_sec} is also often written using the \emph{Riemann symbol}:
%
\begin{linenomath*}
\begin{equation*}
\left \{
\begin{array}{cccc}
0 & 1 & \infty &  \\
\lambda & \mu & \nu & z^+\\
\lambda' & \mu' & \nu' & \\
\end{array}
\right \}
\end{equation*}
\end{linenomath*}
%
where, in this case, $\mu$ and $\mu'$ are as above, $\lambda=\lambda'=0$ and $\nu=0$, $\nu'=2$.
%
To retrieve the general solution to Eq. \ref{can_form} it is sufficient to substitute $(z^++\epsilon^+)/(H^+ +2 \epsilon^+)$ in place of $z^+$ in Eq. \ref{gen_sol_g}.
As the integration constants $\alpha$ and $\beta$ are specified through the imposition of boundary conditions, back-substituting $b^+ = (f_1+f_2)/2$ and $u^+ = -(f_1-f_2)/2i$ yields the solution in terms of $u^+$ and $b^+$.

% ####################################################################

@@

Profiles for N1 are derived by solving monolithically the system of ODEs defined by Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm}, adopting a second-order accurate centered finite difference discretization with over $40,000$ collocation points and stretching the grid via a quadratic coordinate transform (at such resolution profiles are invariant to further refinements down to machine precision).

@@

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{/Users/mg/Phd/papers/bib_files/library}

\begin{thebibliography}{26}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Chu(1987)}]{Chu1987}
Chu PC (1987) {An instability theory of ice-air interaction for the formation
  of ice edge bands}. Journal of Geophysical Research 92(7):6966,
  \doi{10.1029/JC092iC07p06966}

\bibitem[{Egger(1985)}]{Egger1985}
Egger J (1985) {Slope Winds and the Axisymmetric Circulation over Antarctica}.
  Journal of the Atmospheric Sciences 42:1859--1867,
  \doi{10.1175/1520-0469(1985)042<1859:SWATAC>2.0.CO;2}

\bibitem[{Fedorovich and Shapiro(2009{\natexlab{a}})}]{Fedorovich2009d}
Fedorovich E, Shapiro A (2009{\natexlab{a}}) {Structure of numerically
  simulated katabatic and anabatic flows along steep slopes}. Acta Geophysica
  57(4):981--1010, \doi{10.2478/s11600-009-0027-4},
  \urlprefix\url{http://www.springerlink.com/index/10.2478/s11600-009-0027-4}

\bibitem[{Fedorovich and Shapiro(2009{\natexlab{b}})}]{Fedorovich2009}
Fedorovich E, Shapiro A (2009{\natexlab{b}}) {Turbulent natural convection
  along a vertical plate immersed in a stably stratified fluid}. Journal of
  Fluid Mechanics 636:41, \doi{10.1017/S0022112009007757}

\bibitem[{Grisogono and Oerlemans(2001)}]{grisogono2001katabatic}
Grisogono B, Oerlemans J (2001) {Katabatic Flow: Analytic Solution for
  Gradually Varying Eddy Diffusivities}. Journal of the Atmospheric Sciences
  58(21):3349--3354,
  \doi{http://dx.doi.org/10.1175/1520-0469(2001)058<3349:KFASFG>2.0.CO;2}

\bibitem[{Grisogono and Oerlemans(2002)}]{Grisogono2002}
Grisogono B, Oerlemans J (2002) {Justifying the WKB approximation in pure
  katabatic flows}. Tellus, A 54(5):453--462,
  \doi{10.1034/j.1600-0870.2002.201399.x}

\bibitem[{Gutman(1983)}]{gutman_1983}
Gutman LN (1983) {On the theory of the katabatic slope wind}. Tellus, A
  35(3):332--335, \doi{10.3402/tellusa.v35i3.11434}

\bibitem[{Gutman and Malbakhov(1964)}]{Gutman1964}
Gutman LN, Malbakhov VM (1964) {On the Theory of the Katabatic Winds of
  Antarctica}. University of Melbourne, Meteorology Department

\bibitem[{Lykosov and Gutman(1972)}]{Lykosov_1972}
Lykosov VN, Gutman LN (1972) {Turbulent boundary layer above a sloping
  inderlying surface}. Proceedings of the USSR Academy of Sciences 8:799--809,
  \urlprefix\url{papers2://publication/uuid/B2EC0692-4251-4224-8185-A1AE8D9B00E9}

\bibitem[{Mahrt(1998)}]{Mahrt1998}
Mahrt L (1998) {Stratified Atmospheric Boundary Layers and Breakdown of
  Models}. In: Theoretical and Computational Fluid Dynamics, vol~11, pp
  263--279, \doi{10.1007/s001620050093},
  \urlprefix\url{http://link.springer.com/10.1007/s001620050093}

\bibitem[{Monti et~al.(2002)Monti, Fernando, Princevac, Chan, Kowalewski, and
  Pardyjak}]{Monti2002}
Monti P, Fernando HJS, Princevac M, Chan WC, Kowalewski TA, Pardyjak ER (2002)
  {Observations of Flow and Turbulence in the Nocturnal Boundary Layer over a
  Slope}. Journal of the Atmospheric Sciences 59:2513--2534,
  \doi{10.1175/1520-0469(2002)059<2513:OOFATI>2.0.CO;2}

\bibitem[{Morse and Feshbach(1953)}]{Morse1954a}
Morse PM, Feshbach H (1953) {Methods of Theoretical Physics}. McGraw-Hill

>>>>

% ABSTRACT
\begin{abstract}
An exact analytic solution of the steady-state Prandtl model equations is derived, valid for spatially varying eddy diffusivities (O'Briens type) and Prandtl number of unity. 
In this formulation profiles show significant variations in both phase and amplitude of minima-maxima with respect to the classic constant eddy viscosity model and the more recent, approximate, WKB solution. 
The near wall region is characterized by a relatively stronger surface inversion and velocity gradients, the low-level-jet is further displaced toward the wall, and its peak velocity strongly depends on the model parameter, suggesting a tighter coupling between dynamics and thermodynamics. 
%The new solution is used to define vertical resolution requirements in numerical studies based on the Prandtl model, which is of practical interest.
\end{abstract}

\section{Introduction}
Natural convection of turbulent stratified flows over sloping surfaces is ubiquitous in nature and engineering, and is of interest not only as a fundamental problem in itself, but also because of the important role it plays over a broad range of scales and applications. 
On a local scale, for instance, it governs local weather conditions, affecting atmospheric transport processes and dispersion of scalars such as heat and humidity \citep{Monti2002, Nylen2004b}. 
On a regional scale, it drives the development of deep-convective circulations \citep{Egger1985, Parish1998} and is responsible for intense cyclonic vorticity in the middle and upper troposphere \citep{Parish1992}.
Further, recent experimental field campaigns \citep{Chu1987, Oerlemans1993} have shown that persistent katabatic winds characterize the atmospheric boundary layer over snow-ice surfaces and glaciers, and therefore an accurate characterization of such flows is an essential component of understanding and modeling of the weather and climate.
However, the complex dynamics (e.g. occurrence of intermittency, waves, Kelvin-Helmholtz instabilities and low-level jets) and the lack of a satisfactory similarity theory for such flows \citep{Nadeau2012} pose a heavy burden in terms of computational requirements for numerical modellers; in most cases the required resolution is prohibitively costly \citep{Fedorovich2009, Fedorovich2009d, Burkholder2011a}.
Because of this, conceptual models are still of great interest, and represent a valid tool for the characterization of such systems. 
A cornerstone in the understanding of natural convection of stratified fluid along a sloping surface is represented by the classic Prandtl analytic model \citep{Prandtl}, and its more recent extensions/generalizations to include the effects of Coriolis force \citep{Gutman1964}, external winds \citep{Lykosov_1972}, surface heterogeneity \citep{Shapiro2007, Oldroyd2014}.
The Prandtl model approximates the atmosphere in a Boussinesq sense and describes a steady flow over a thermally perturbed unbounded planar sloping surface that lies within a stratified environment, given by the following system of equations:
%
\begin{linenomath*}
\begin{eqnarray}
	N^2 u(z) \sin{\alpha} & = & d( K_H db/dz )/dz,
	\label{eqn1u} \\
	%
	b(z) \sin{\alpha} & = & -d( K_M du/dz )/dz,
	\label{eqn1b}
\end{eqnarray}
\end{linenomath*}
%
where $z$ denotes the normal-to-slope coordinate directions, $u$ is the downslope velocity, $b$ is buoyancy, $N$ is the buoyancy frequency characterizing the system (related to the background stratification), $\alpha$ is the slope angle and $K_M$ and $K_H$ denote the eddy viscosity and diffusivity (an eddy viscosity/diffusivity model has been used to parametrize turbulent fluxes of momentum and buoyancy).
Equations are defined in $z \in \ [0,\infty)$ with boundary conditions $u(0)=0$, $u(z \rightarrow \infty) = 0$, $b(0)=b_s$ and $b(z \rightarrow \infty) = 0$ ($b_s > 0$ for upslope flows, whereas $b_s<0$ for downslope flows).

The flow is assumed to be invariant in the downslope direction and the model can be used to determine its vertical structure.
Eqs. \ref{eqn1u} and \ref{eqn1b} state that downslope (upslope) convection of cold (warm) air is balanced by momentum flux divergence, and that along-slope generated buoyancy is balanced by buoyancy flux divergence; thus the model is applicable away from ridges and valleys, so that advection terms are negligible \citep{NAPPO1987}.
The Prandtl constant-K model is known to be overdissipative in the near surface regions, and therefore not able to represent the observed strong surface gradients of temperature and momentum that commonly characterize katabatic and anabatic flows \citep{Oerlemans1998, grisogono2001katabatic}. 
Simple variations in the eddy diffusivity profiles were introduced in a matched analytic solution by \citet{gutman_1983}, and more recently \citet{grisogono2001katabatic} considered general variations in the vertical structure of the eddy diffusivities, and derived  solutions based on the WKB approximation \citep{Bender1999}.
The WKB solution is able to account for additional dynamics while still retaining an elegant form.
However, that theory is only applicable when the model parameters $(K_M, K_H)$ vary more slowly than the solution $(u,b)$, and the validity of such a constraint for slope flows has been the subject of great debate \citep{Grisogono2002}.
In the following, we derive a closed-form solution to the Prandtl-model equations, valid for eddy viscosity / diffusivity coefficients that are modelled as a limited range of cubic polynomials, similar to what proposed in \citet{O'Brien1970} for the planetary boundary layer. 
The solution thus represents an exact alternative to the WKB formulation for the chosen form of the model parameter.
The new solution, expressed as a combination of Gaussian hypergeometric functions, is found to significantly differ from its analytical counterparts, shedding new light on the problem, and suggesting a different coupling between the dynamic and the thermodynamic boundary layers.

% ###################################################################
\section{The Analytic Solution}
%
We consider the one-dimensional Prandtl model (Eq. \ref{eqn1u} and \ref{eqn1b}).

Assuming eddy diffusivities in line with the classic O'Brien's type \citep{O'Brien1970}, viz. $K(z) = A (z + \epsilon)(z-\zeta)^2$, where $\epsilon>0$ and $\zeta>H$ are parameters that we introduce in order to shift the zeroes of $K$ outside the interval of integration $[0,H]$, assuming $Pr=1$ (i.e. $K_M=K_H=K$) and assigning a length, velocity and buoyancy scale $L= (N \sin{\alpha})/ A$, $U=b_s N^{-1}$ and $B=b_s$ respectively, Eqs. \ref{eqn1u} and \ref{eqn1b} reduce to:
\begin{linenomath*}
\begin{subequations}
\label{ode}
\begin{align}
	\overline{u} & = d( \overline{K} d\overline{b}/d\overline{z} )/d\overline{z}, && \mathrm{in}  \ [0,\overline{H}]
	\label{eqn1u_norm} \\
	\overline{b} & = -d( \overline{K} d\overline{u}/d\overline{z} )/d\overline{z}, &&\mathrm{in}  \ [0,\overline{H}]
	\label{eqn1b_norm}
\end{align}
\end{subequations}
\end{linenomath*}
where $\overline{K} = (\overline{z} + \overline{\epsilon})(\overline{z}-\overline{\zeta})^2$ is the normalized eddy viscosity/diffusivity, $\overline{z}=zL^{-1}$, $\overline{u} = uU^{-1}$, $\overline{b}=b B^{-1}$, with boundary conditions $\overline{u}(0)=0$, $\overline{u}(\overline{H}) = 0$, $\overline{b}(0)=\pm 1$ and $\overline{b}(\overline{H}) = 0$.
Overbars $\overline{(\cdot)}$ will be omitted hereafter.
Further, we introduce the following variable
\begin{linenomath*}
\begin{equation}
   f= b + i u,
   \label{combinationVar}
\end{equation}
\end{linenomath*}
and upon substitution of Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm} into Eq. \ref{combinationVar}, the system results into a complex ordinary differential equation (ODE) for the canonical variable $f$:
\begin{linenomath*}
\label{ode_norm}
\begin{equation}
	i \cdot f =  d(  K df/dz )/dz \hspace{1cm} \mathrm{in}  \ [0,H],
	\label{ode}
\end{equation}
\end{linenomath*}
with boundary conditions $f(0)=\pm 1$ and $f(H) = 0$.
This technique for the decoupling of the system of Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm} is based on an ad hoc diagonalization.
To solve Eq. \ref{ode} we first rewrite the equation in canonical form:
\begin{linenomath*}
\begin{equation}
	\label{can_form}
	f''+P f'+ Qf =0,
\end{equation}
\end{linenomath*}
where $(\cdot)' = \frac{\partial (\cdot)}{\partial z}$, $P(z)=(K)'/K$ and $Q(z)=1/(i K)$. Second, it is an easy computation to show that rewriting Eq. \ref{can_form} for $y = \frac{z + \epsilon}{H + \epsilon + \zeta}
$ results in
%if $f$ solves Eq. \ref{can_form} then $g(z^+):=f( (H^++2 \epsilon^+) z^+- \epsilon^+)$ solves:
\begin{linenomath*}
\begin{equation}
	\label{can_form_sec}
	f''+\widetilde{P} f'+ \widetilde{Q} f =0,
\end{equation}
\end{linenomath*}
where $(\cdot)' = \frac{\partial (\cdot)}{\partial y}$, $\widetilde{P}(y)=\gamma'(y)/\gamma(y)$ and $\widetilde{Q}(y)=1/( i \cdot (H+\epsilon+\zeta) \gamma(y))$, with $\gamma(y)=y(y-1)^2$.
Eq. \ref{can_form_sec} is a second order ODE with three regular singular points at $y = 0,1$ and $\infty$, as in \citet{Morse1954a}. 
This special case is known as the \emph{equation of Papperitz} and its general solution is:
%\begin{widetext}
\begin{linenomath*}
\begin{equation}
	f(y) = \alpha (1-y)^\mu \ _2F_1 (\mu, 1- \mu', 1+\mu-\mu', 1-y)
	+\beta (1-y)^{\mu'} \ _2F_1 (\mu', 1- \mu, 1+\mu'-\mu, 1-y),
	\label{gen_sol_g}
\end{equation}
\end{linenomath*}
%\end{widetext}
where $\ _2F_1$ are \emph{Gaussian hypergeometric functions} and $\mu$, $\mu'$ are the solutions to the degree two equation
\begin{linenomath*}
\begin{equation*}
	x^2+x+1/( i (H+ \epsilon + \zeta))=0.
\end{equation*}
\end{linenomath*}
Note that the general solution to Eq. \ref{can_form_sec} is also often written using the \emph{Riemann symbol}:
\begin{linenomath*}
\begin{equation*}
\left \{
\begin{array}{cccc}
0 & 1 & \infty &  \\
\lambda & \mu & \nu & y\\
\lambda' & \mu' & \nu' & \\
\end{array}
\right \}
\end{equation*}
\end{linenomath*}
where, in this case, $\mu$ and $\mu'$ are as above, $\lambda=\lambda'=0$ and $\nu=0$, $\nu'=2$.
Upon back-substitution of the independent variable and specification of the integration constants $\alpha$ and $\beta$ (through the imposition of boundary conditions), the solution in terms of $u$ and $b$ is derived by separating the real and imaginary part of $f$
\begin{equation}
	u(z) = \operatorname{Re}(f(z)), \hspace{1cm} b(z) = \operatorname{Im}(f(z)).
\end{equation}

% ####################################################################
\section{Specification of of K(z)}

The model assumes $Pr = 1$, i.e. $K_M = K_H = K$, although in reality they are slightly different. This assumption is justified based on the degree of accuracy of the considered study. 
Further, as stated in the preceeding section, we assume eddy diffusivities in line with the classic O'Brien's type \citep{O'Brien1970}, viz. $K = A(z + \epsilon)(z-\zeta)^2$, where $\epsilon>0$ can be tuned to provide the desired value of eddy viscosity / diffusivity at the wall, whereas $\zeta > H$ controls both shape and magnitude of $K$.
The O'Brien model has often been used in studies of stable boundary layers, see for instance \citet{pielke1984mesoscale} and \citet{Stull1988}, and a generalized O'Brien model was also recently adopted in \citet{grisogono2001katabatic} to study slope flows.
The properties of $K(z)$ in slope flows were thorougly justified in \citet{Grisogono2002};  we here recall the main criteria that $K(z)$ should honor:
%
\begin{enumerate}
\item realizability condition: $K(z) \ge 0$;
\item $K(z << \infty) = 0$;
\item $\max(K(z))$, reached at $z=h$, must lie within the dynamic and thermal boundary layers;
\item $h > \max(2 z_j, z_{inv})$;
\end{enumerate}
%
where $z_j$ is the height of the LLJ, and $z_{inv}$ is the height of the surface inversion layer (in general we have $z_{inv} > z_j$).
%
In addition, assuming Monin-Obukhov similarity theory to hold for the near wall regions (which is itself a subject of great debate), one would ideally specify $K(z)$ in order to satisfy:
%
\begin{linenomath*}
\begin{subequations}
\begin{align}
   K(z) & = \kappa u_* z/[1+\Phi(z)] & z \in [z_0,z_s], \label{MOST1}\\
   K(z) & = f(u_*,z) \ | \ d^2f/dz^2 < 0 & z \in (z_s,H], \label{MOST2}
\end{align}
\end{subequations}
\end{linenomath*}
%
where $z_s$ is the upper boundary of the constant-flux layer, $\Phi(z)$ is a stability correction function, $\kappa$ is the Von K\'arm\'an constant, and $u_* = \sqrt{K(du/dz)/\rho}$ is the so-called friction velocity.
However, in the current settings, we are not able to satisfy Eqs. \ref{MOST1} and \ref{MOST2} directly, since the specification of $K$ has to be done a-priori, and $u_*$ depends on the solution. 
This is the main weakness of linear approaches. The only way to enforce Eqs. \ref{MOST1} and \ref{MOST2} would be to apply a procedure in which $K(z)$ is iteratively updated to match the MO similarity relations.
Based on recent direct (DNS) and large-eddy simulation (LES) results (not shown) we make the additional assumption that $h$ is located in the neighborhood of the first zero of the $u$ velocity profile, well above $z_j$ but within the dynamic and thermal boundary layers ($h$ would be at the lower boundary of the return flow).
Knowledge of $K(z_0),dK(z_0)/dz$ and $h$ allows to univocally define $K(z)= A(z + \epsilon)(z-\zeta)^2$ one can therefore specify the surface value and its derivative $K(0), dK(0)/dz$ and 
Under such restrictive (but realistic) assumption it is possible to show that the $H$ parameter in

Despite the restrictive form of the model parameter, which might limit the applicability of the solution in certain contexts, the model

@@

Profiles for N1 are derived as direct numerical solution the system of ODEs defined by Eqs. \ref{eqn1u_norm} and \ref{eqn1b_norm}, where a second-order accurate centered finite difference discretization is adopted with over $40,000$ collocation points and stretching the grid via a quadratic coordinate transform (at such resolution profiles are invariant to further refinements down to machine precision).

@@

% BibTeX users please use one of
\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{/Users/mg/Phd/papers/bib_files/library}

%\begin{thebibliography}{26}
%\providecommand{\natexlab}[1]{#1}
%\providecommand{\url}[1]{{#1}}
%\providecommand{\urlprefix}{URL }
%\expandafter\ifx\csname urlstyle\endcsname\relax
%  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
%  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
%  \urlstyle{rm}\Url}\fi
%\providecommand{\eprint}[2][]{\url{#2}}
%\bibitem[{Bender and Orszag(1999)}]{Bender1999}
%Bender CM, Orszag SA (1999) {Advanced Mathematical Methods for Scientists and
%  Engineers I}. Springer New York, New York, NY,
%  \doi{10.1007/978-1-4757-3069-2},
%  \urlprefix\url{http://eprints.ucl.ac.uk/70187/
%  http://link.springer.com/10.1007/978-1-4757-3069-2}
%\bibitem[{Chu(1987)}]{Chu1987}
%Chu PC (1987) {An instability theory of ice-air interaction for the formation
%  of ice edge bands}. Journal of Geophysical Research 92(7):6966,
%  \doi{10.1029/JC092iC07p06966}
%\bibitem[{Egger(1985)}]{Egger1985}
%Egger J (1985) {Slope Winds and the Axisymmetric Circulation over Antarctica}.
%  Journal of the Atmospheric Sciences 42:1859--1867,
%  \doi{10.1175/1520-0469(1985)042<1859:SWATAC>2.0.CO;2}
%\bibitem[{Fedorovich and Shapiro(2009{\natexlab{a}})}]{Fedorovich2009d}
%Fedorovich E, Shapiro A (2009{\natexlab{a}}) {Structure of numerically
%  simulated katabatic and anabatic flows along steep slopes}. Acta Geophysica
%  57(4):981--1010, \doi{10.2478/s11600-009-0027-4},
%  \urlprefix\url{http://www.springerlink.com/index/10.2478/s11600-009-0027-4}
%\bibitem[{Fedorovich and Shapiro(2009{\natexlab{b}})}]{Fedorovich2009}
%Fedorovich E, Shapiro A (2009{\natexlab{b}}) {Turbulent natural convection
%  along a vertical plate immersed in a stably stratified fluid}. Journal of
%  Fluid Mechanics 636:41, \doi{10.1017/S0022112009007757}
%\bibitem[{Grisogono and Oerlemans(2001)}]{grisogono2001katabatic}
%Grisogono B, Oerlemans J (2001) {Katabatic Flow: Analytic Solution for
%  Gradually Varying Eddy Diffusivities}. Journal of the Atmospheric Sciences
%  58(21):3349--3354,
%  \doi{http://dx.doi.org/10.1175/1520-0469(2001)058<3349:KFASFG>2.0.CO;2}
%\bibitem[{Grisogono and Oerlemans(2002)}]{Grisogono2002}
%Grisogono B, Oerlemans J (2002) {Justifying the WKB approximation in pure
%  katabatic flows}. Tellus, A 54(5):453--462,
%  \doi{10.1034/j.1600-0870.2002.201399.x}
%\bibitem[{Gutman(1983)}]{gutman_1983}
%Gutman LN (1983) {On the theory of the katabatic slope wind}. Tellus, A
%  35(3):332--335, \doi{10.3402/tellusa.v35i3.11434}
%\bibitem[{Gutman and Malbakhov(1964)}]{Gutman1964}
%Gutman LN, Malbakhov VM (1964) {On the Theory of the Katabatic Winds of
%  Antarctica}. University of Melbourne, Meteorology Department
%\bibitem[{Lykosov and Gutman(1972)}]{Lykosov_1972}
%Lykosov VN, Gutman LN (1972) {Turbulent boundary layer above a sloping
%  inderlying surface}. Proceedings of the USSR Academy of Sciences 8:799--809,
%  \urlprefix\url{papers2://publication/uuid/B2EC0692-4251-4224-8185-A1AE8D9B00E9}
%\bibitem[{Mahrt(1998)}]{Mahrt1998}
%Mahrt L (1998) {Stratified Atmospheric Boundary Layers and Breakdown of
%  Models}. In: Theoretical and Computational Fluid Dynamics, vol~11, pp
%  263--279, \doi{10.1007/s001620050093},
%  \urlprefix\url{http://link.springer.com/10.1007/s001620050093}
%\bibitem[{Monti et~al.(2002)Monti, Fernando, Princevac, Chan, Kowalewski, and
%  Pardyjak}]{Monti2002}
%Monti P, Fernando HJS, Princevac M, Chan WC, Kowalewski TA, Pardyjak ER (2002)
%  {Observations of Flow and Turbulence in the Nocturnal Boundary Layer over a
%  Slope}. Journal of the Atmospheric Sciences 59:2513--2534,
%  \doi{10.1175/1520-0469(2002)059<2513:OOFATI>2.0.CO;2}
%\bibitem[{Morse and Feshbach(1953)}]{Morse1954a}
%Morse PM, Feshbach H (1953) {Methods of Theoretical Physics}. McGraw-Hill
%\bibitem[{Nadeau et~al.(2013)Nadeau, Pardyjak, Higgins, and
%  Parlange}]{Nadeau2012}
